On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-05 arity1.tex
Add files via upload
Succinctness, succinctness

	\end{minipage}
	\caption{}
	\label{table:transcript}
\end{figure}

@@

In subsections~\ref{sec:unary}.1 and \ref{sec:unary}.2, we will argue that unary systems at arity $1$ and delay $2$ and $3$, respectively, cannot fold infinite transcripts deterministically. As we will see, in fact, the length of transcripts deterministically foldable by these systems has an upper bound quadratic in the length of the seed. The main result is the following theorem, which is a direct consequence of Lemma x and Lemma y from the next two subsections.
\begin{theorem}[$\delta\in\{ 2,3\}, \alpha = 1$]\label{thm:d23a1}
	Let $\Xi$ be a unary oritatami system of $\delta \in \{2,3\}$ and $\alpha = 1$. Starting with a seed of length $n$, $\Xi$ can yield only finite structures of size at most $3n^2+4n+2$.
\end{theorem}

@@

\subsubsection{\ref{sec:unary}.1 \ \ $\delta = 2$\\}
\begin{lemma}[$\delta=2, \alpha =1$]\label{lem:d2a1}
	Let $\Xi$ be a unary oritatami system of $\delta=2$ and $\alpha = 1$. Starting with a seed of length $n$, $\Xi$ can yield only finite structures of size at most $3n^2+4n+2$.
\end{lemma}
\begin{proof}

@@

\noindent\textbf{Case 2.} bead $i+1$ is fixed at $(0,1)$. This is only possible if
	\begin{enumerate}[label=(\alph*)]
		\item there is an inactive bead at $(-1,0)$ and one with a free hand one at $(-2,0)$. This case is symmetrical to (1).
		\item there is no bead at $(-1,0)$, bead $i+1$ can bond with bead $i-1$ at $(0,0)$ and the bead $i+2$ can be placed at $(-1,0)$ where it can bond with $(-2,0)$, $(-2,-1)$ or $(-1,-1)$. This leads to nondeterminism, because placing bead $i$ at $(-1,0)$ and bead $i+1$ at $(0,1)$ would yield two bonds, just as the original conformation.
		\item there is a bead at $(-1,0)$ and bead $i+1$ can bond with that or with bead $i-1$ at $(0,0)$. However, this means that placing bead $i$ at $(0,1)$ at bead $i+1$ at $(1,1)$ creates the same number of hydrogen bonds, thus resulting in bead $i$ not being placed deterministically.

	\end{enumerate}
%\end{enumerate}

@@

\subsubsection{\ref{sec:unary}.2 \ \ $\delta = 3$\\}

@@

As before, let us assume that the last bead stabilized within $\hexagon_p^n$ is at point $(0,0)$ and the next bead is the first to be stabilized outside $\hexagon_p^n$ at point $(1,1)$. Depending on whether bead $i-1$ at point (0,0) has a free hand to form a bond or not, we distinguish two cases.\\
%\begin{enumerate}
\noindent\textbf{Case 1.}  Bead $i-1$ at (0,0) has a free hand.\\ If the most stable conformation formed by beads $i, i+1, i+2$ adds only two  bonds, it will be nondeterministic because there are at least two possibilities (see Fig.~\ref{fig:beadi2}, except it starts from (0,0) not (1,1)). Therefore, it needs to make three new bonds to deterministically stabilize $i$.\\

%\begin{figure}
%  \begin{center}
%    \begin{tabular}{cc}
%      \begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
%        \fill (0.0, 0.0) circle [radius = 0.1];
%        \fill (0.5, 0.866) circle [radius = 0.1];
%        \fill (-0.5, 0.866) circle [radius = 0.1];
%        \fill (0.0, 1.732) circle [radius = 0.1];
%        \fill (0.0, 0.0) node [below] {$(0, 0)$};
%        \draw (0.0, 0.0) -- (0.5, 0.866);
%        \draw (0.5, 0.866) -- (-0.5, 0.866);
%        \draw (0.0, 1.732) -- (-0.5, 0.866);
%        \draw [dashed] [red] (0.0, 1.732) -- (0.5, 0.866);
%        \draw [dashed] [red] (0.0, 0.0) -- (-0.5, 0.866);
%      \end{tikzpicture}
%
%      \begin{tikzpicture}[thick,scale=0.8, every node/.style={scale=0.8}]
%        \fill (0.0, 0.0) circle [radius = 0.1];
%        \fill (-0.5, 0.866) circle [radius = 0.1];
%        \fill (0.5, 0.866) circle [radius = 0.1];
%        \fill (0.0, 1.732) circle [radius = 0.1];
%        \fill (0.0, 0.0) node [below] {$(0, 0)$};
%        \draw (0.0, 0.0) -- (-0.5, 0.866);
%        \draw (-0.5, 0.866) -- (0.5, 0.866);
%        \draw (0.0, 1.732) -- (0.5, 0.866);
%        \draw [dashed] [red] (0.0, 0.0) -- (0.5, 0.866);
%        \draw [dashed] [red] (0.0, 1.732) -- (-0.5, 0.866);
%      \end{tikzpicture}
%
%    \end{tabular}
%    \caption{Two conformations when bead $i-1$ at (0,0) has a free hand}
%    \label{fig:beadi}
%  \end{center}
%\end{figure}

@@

There are five possible cases in which beads $i, i+1, i+2$ can add three bonds, see Fig.~\ref{fig:3bonds}. In the cases (b) and (e) in Fig.~\ref{fig:3bonds}, there are beads having a free hand at (1,0) and (-1,0) already stabilized before bead $i-1$ is fixed at (0,0). One of these two beads may be a predecessor of bead $i-1$, but at least one of them is not. When bead $i-1$ is fixed at (0,0), it makes a bond with the one of these two beads which is not its predecessor. This means that it is impossible to have three beads at (-1,0), (0,0) and (1,0), each with a free hand, when bead $i$ is fixed, and consequently, cases (b) and (e) in Fig.~\ref{fig:3bonds} cannot occur. Case (d) in Fig.~\ref{fig:3bonds} becomes nondeterministic when bead $i$ is fixed because bead $i$ can be fixed at (-1,0) and bond with (-2,0), bead $i+1$ can be placed (0,1) and bond with (0,0) and bead $i+2$ can be placed (1,1) and bond with (0,1). If it makes three bonds once, such as (a) and (c) in Fig.~\ref{fig:3bonds}, it will need to make three bonds forever to be deterministic. Similarly to case 1. in the previous section, cases (a) and (c) in Fig.~\ref{fig:3bonds} lead to nondeterministism eventually, when the transcript reaches a first corner of $\hexagon_p^n$.

@@

%There are not any bead having a free hand at (1,0) or (-1,0) to prevent from being (a) and (c) in Fig.~\ref{fig:3bonds}. 
(ii) Now consider when there is no bead at (1,0) or there is one, but with no free hand. Let us discuss the moment after bead $i$ is fixed outside $\hexagon_p^n$. Now bead $i$ has a free hand because it cannot bind to (1,0). If beads $i+1, i+2, i+3$ can form only two bonds, it will be nondeterministic because there are at least two possible such conformations, as in Fig.~\ref{fig:beadi2}. Hence, they need to form three bonds to deterministically stabilize, such as in Fig.~\ref{fig:3bonds2}. Case (a) in Fig.~\ref{fig:3bonds2}. becomes the same as case (a) in Fig.~\ref{fig:3bonds}. Case (b) in Fig.~\ref{fig:3bonds2} becomes nondeterministic already when bead $i$ is fixed because bead $i$ could also be fixed at (1,0).\\

@@

    \end{tabular}
    \caption{When beads $i+1, i+2, i+3$ make three hydrogen bonds (a) there is an inactive bead at (1,0). (b) there is no bead at (1,0).}
    \label{fig:3bonds2}
  \end{center}
\end{figure}

\qed
\end{proof}

We have shown that at $\alpha=1$ and  $\delta\in\{2,3\}$, unary oritatami systems can only fold finite length transcripts deterministically, and the length of that transcript is bounded by the size of the regular hexagon of radius $n$ plus the length of one side of the surrounding hexagon. This gives us the upper bound $|\hexagon_p^{n}|+(n+1)=3n^2+4n+2$,  where $n$ is the length of the seed, concluding the proof of Theorem~\ref{thm:d23a1}.

>>>>

	\end{minipage}
	\caption{Transcript of pyramid-like conformation at delay $2$, arity $1$.}
	\label{table:transcript}
\end{figure}

@@

In this section we will argue that unary systems at arity $1$ and delay $2$ and $3$, respectively, cannot fold infinite transcripts deterministically. As we will see, in fact, the length of transcripts deterministically foldable by these systems has an upper bound quadratic in the length of the seed. The main result is the following theorem, which is a direct consequence of Lemma~\ref{lem:d2a1}  and Lemma~\ref{lem:d3a1} which follow.
\begin{theorem}[$\delta\in\{ 2,3\}, \alpha = 1$]\label{thm:d23a1}
	Let $\Xi$ be a unary oritatami system of $\delta \in \{2,3\}$ and $\alpha = 1$. Starting with a seed of length $n$, $\Xi$ can yield only finite structures of size at most $3n^2+4n+2$.
\end{theorem}

@@

%\subsubsection{\ref{sec:unary}.1 \ \ $\delta = 2$\\}
\begin{lemma}[$\delta=2, \alpha =1$]\label{lem:d2a1}
	Let $\Xi$ be a unary oritatami system of $\delta=2$ and $\alpha = 1$. Starting with a seed of length $n$, $\Xi$ can yield only finite structures of size at most $3n^2+4n+2$.
\end{lemma}
\begin{proof}

@@

\noindent\textbf{Case 2.} bead $i+1$ is fixed at $(0,1)$. This is only possible if\\
\textbf{(a)} there is an inactive bead at $(-1,0)$ and one with a free hand one at $(-2,0)$. This case is symmetrical to (1).
there is no bead at $(-1,0)$, bead $i+1$ can bond with bead $i-1$ at $(0,0)$ and the bead $i+2$ can be placed at $(-1,0)$ where it can bond with $(-2,0)$, $(-2,-1)$ or $(-1,-1)$. This leads to nondeterminism, because placing bead $i$ at $(-1,0)$ and bead $i+1$ at $(0,1)$ would yield two bonds, just as the original conformation.\\
\textbf{(b)	} there is a bead at $(-1,0)$ and bead $i+1$ can bond with that or with bead $i-1$ at $(0,0)$. However, this means that placing bead $i$ at $(0,1)$ at bead $i+1$ at $(1,1)$ creates the same number of hydrogen bonds, thus resulting in bead $i$ not being placed deterministically.

%\end{enumerate}

@@

%\subsubsection{\ref{sec:unary}.2 \ \ $\delta = 3$\\}

@@

As before, let us assume that the last bead stabilized within $\hexagon_p^n$ is at point $(0,0)$ and the next bead is the first to be stabilized outside $\hexagon_p^n$ at point $(1,1)$. Depending on whether bead $i-1$ at point (0,0) has a free hand to form a bond or not, we distinguish two cases.\\
%\begin{enumerate}
\noindent\textbf{Case 1.}  Bead $i-1$ at (0,0) has a free hand. If the most stable conformation formed by beads $i, i+1, i+2$ adds only two  bonds, it will be nondeterministic because there are at least two possibilities (see Fig.~\ref{fig:beadi2}, except it starts from (0,0) not (1,1)). Therefore, it needs to make three new bonds to deterministically stabilize $i$. There are five possible cases in which beads $i, i+1, i+2$ can add three bonds, see Fig.~\ref{fig:3bonds}. In the cases (b) and (e) in Fig.~\ref{fig:3bonds}, there are beads having a free hand at (1,0) and (-1,0) already stabilized before bead $i-1$ is fixed at (0,0). One of these two beads may be a predecessor of bead $i-1$, but at least one of them is not. When bead $i-1$ is fixed at (0,0), it makes a bond with the one of these two beads which is not its predecessor. This means that it is impossible to have three beads at (-1,0), (0,0) and (1,0), each with a free hand, when bead $i$ is fixed, and consequently, cases (b) and (e) in Fig.~\ref{fig:3bonds} cannot occur. Case (d) in Fig.~\ref{fig:3bonds} becomes nondeterministic when bead $i$ is fixed because bead $i$ can be fixed at (-1,0) and bond with (-2,0), bead $i+1$ can be placed (0,1) and bond with (0,0) and bead $i+2$ can be placed (1,1) and bond with (0,1). If it makes three bonds once, such as (a) and (c) in Fig.~\ref{fig:3bonds}, it will need to make three bonds forever to be deterministic. Similarly to case 1. in the previous section, cases (a) and (c) in Fig.~\ref{fig:3bonds} lead to nondeterministism eventually, when the transcript reaches a first corner of $\hexagon_p^n$.

@@

@@

%There are not any bead having a free hand at (1,0) or (-1,0) to prevent from being (a) and (c) in Fig.~\ref{fig:3bonds}. 
\noindent (ii) Now consider when there is no bead at (1,0) or there is one, but with no free hand. Let us discuss the moment after bead $i$ is fixed outside $\hexagon_p^n$. Now bead $i$ has a free hand because it cannot bind to (1,0). If beads $i+1, i+2, i+3$ can form only two bonds, it will be nondeterministic because there are at least two possible such conformations, as in Fig.~\ref{fig:beadi2}. Hence, they need to form three bonds to deterministically stabilize, such as in Fig.~\ref{fig:3bonds2}. Case (a) in Fig.~\ref{fig:3bonds2}. becomes the same as case (a) in Fig.~\ref{fig:3bonds}. Case (b) in Fig.~\ref{fig:3bonds2} becomes nondeterministic already when bead $i$ is fixed because bead $i$ could also be fixed at (1,0).\\

@@

    \end{tabular}
    \caption{When beads $i+1, i+2, i+3$ make three hydrogen bonds (a) there is an inactive bead at (1,0). (b) there is no bead at (1,0).}
    \label{fig:3bonds2}
  \end{center}
\end{figure}
We have shown that at $\alpha=1$ and  $\delta\in\{2,3\}$, unary oritatami systems can only fold finite length transcripts deterministically, and the length of that transcript is bounded by the size of the regular hexagon of radius $n$ plus the length of one side of the surrounding hexagon. This gives us the upper bound $|\hexagon_p^{n}|+(n+1)=3n^2+4n+2$,  where $n$ is the length of the seed, concluding the proof of Theorem~\ref{thm:d23a1}.
\end{proof}
